Open Data, Open Source and Open Standards in chemistry: \\ The Blue Obelisk five years on

### 2011-06-09 paper.tex
Done for today.

>>>>

\subsection*{Reaction Attempts}
Although there are existing databases that allow for searching reactions, those
using Open Data are harder to find. The Reaction Attempts database
\cite{ReactionAttempts}, to which anyone can
submit reaction attempts data, consists mainly of reaction information
abstracted from Open Notebooks in organic chemistry, such as the
UsefulChem project from the Bradley group
\cite{UsefulChemReactionAttempts}
and the notebooks from the Todd group \cite{UsefulChemTodd}.
Key information from each experiment is abstracted manually, with the
only required information consisting of the ChemSpider IDs of the
reactants and the product targeted in the experiment; and a link to
the laboratory notebook page.  Since the database reflects all data
from the notebooks, many experiments will be short of completely
optimized and includes experiments in progress, ambiguous results and
failed runs.  Unlike most reaction databases that only identify
experiments successfully reported in the literature, Reaction Attempts
is meant to be used to easily discover what researchers are trying to
accomplish thus potentially forging collaborations more quickly~\cite{UsefulChemSocialNetworks}
or providing useful knowledge about what doesn't work.

The database can be accessed using
a web-based Reaction Attempts Explorer
\cite{ReactionAttemptsExplorer}.
The explorer allows you to find reaction attempts either by reactant
or product name, via a series of dynamic dropdowns, or by performing a
substructure search. Found reactions are shown graphically as schemes
with accompanying structure images and text with links to the lab
notebook pages, the ChemSpider pages for each compound, and a separate
web service, The Optimal Solvent Selector
[http://usefulchem.blogspot.com/2010/08/reaction-attempts-solvent-selector.html,
that predicts the solubility of the reactants and product in 72
solvents and sorts by theoretical maximum yield.  The Reaction
Attempts Explorer allows researchers to easily find patterns in
reactions that have already been performed and since the data are open
and results are reported across all research groups, intersections are
easily discovered and possible Open Collaboration opportunities are
easily found \cite{ekins_collaboration_2011}.
